---
title: "Classification Intro"
author: "Bastola"
date: "`r format(Sys.Date(), ' %B %d %Y')`"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      collapse = TRUE, 
                      comment=NA, 
                      warning = FALSE,
                      message = FALSE,
                      fig.height = 4, fig.width = 6, fig.align='center')

library(tidyverse) 
library(caret)
library(class)
library(patchwork)
library(tidymodels)
library(mlbench)     # for PimaIndiansDiabetes2 dataset
library(janitor)
library(parsnip)
library(kknn)
library(paletteer)
library(corrr)
library(scico)
library(forcats)
library(yardstick)

select <- dplyr::select

# Load the fire data

fire <- read_csv("https://raw.githubusercontent.com/deepbas/statdatasets/main/Algeriafires.csv")
fire <- fire %>% clean_names() %>% na.omit() %>% mutate_at(c(10,13), as.numeric)
#fire1 <- fire %>% mutate(across(where(is.numeric), standardize))

fire_raw <- fire %>% dplyr::select(temperature, isi, classes)

fire_recipe <- recipe(classes ~ ., data = fire_raw) %>%
 step_scale(all_predictors()) %>%
 step_center(all_predictors()) %>%
 prep()

fire_scaled <- bake(fire_recipe, fire_raw)

fire_knn_spec <- nearest_neighbor(mode = "classification",
                             engine = "kknn",
                             weight_func = "rectangular",
                             neighbors = 5)

fire_knn_fit <- fire_knn_spec %>%
 fit(classes ~ ., data = fire_scaled)

fire_knn_workflow <- workflow() %>% 
  add_recipe(fire_recipe) %>%
  add_model(fire_knn_spec)

fire_fit <- fit(fire_knn_workflow, data = fire_raw)

```

## Your Turn 1

1. Load the `mlbench` package to get `PimaIndiansDiabetes2` dataset.

```{r}
# Load the data - diabetes
data(PimaIndiansDiabetes2)
db <- PimaIndiansDiabetes2
db <- db %>% na.omit() %>% mutate(diabetes = fct_rev(factor(diabetes))) 
db_raw <- db %>% select(glucose, insulin, diabetes)

```


a. Set aside 20% of the cases using the following code. We'll make predictions on those penguins later.

```{r}
set.seed(124544656)

db_split <- initial_split(db, prop = 0.50)

# Create training data
db_train <- db_split %>%
                    training()

# Create testing data
db_test <- db_split %>%
                    testing()

```
  
  
b. Follow the steps to train a 7-NN classifier using the `tidymodels` toolkit 
    

```{r}
db_recipe <- recipe(diabetes ~ ., data = db_raw) %>%
  step_scale(all_predictors()) %>%
  step_center(all_predictors()) %>%
  prep()
```
    

```{r}
db_knn_spec7 <- nearest_neighbor(mode = "classification",
                             engine = "kknn",
                             weight_func = "rectangular",
                             neighbors = 7)

db_workflow <- workflow() %>% 
  add_recipe(db_recipe) %>%
  add_model(db_knn_spec7)
```


```{r}
db_fit <- fit(db_workflow, data = db_train)
```
    
c. Classify the penguins in the `test` data frame. How did you do?

```{r}
test_features <- db_test %>% select(glucose, insulin) %>% data.frame()
db_pred <- predict(db_fit, test_features, type = "raw")

db_results <- db_test %>% 
  select(diabetes) %>% 
  bind_cols(predicted = db_pred)

# Accuracy

(accuracy <- sum(which(db_results$diabetes == db_results$predicted))/ nrow(db_results))

```


3. Repeat the above steps with 50-50 split in the dataset.


# Your Turn 2


```{r}


```




```{r echo=FALSE}
set.seed(234)

fire_split3 <- initial_split(fire_raw, prop = 0.6)

# Create training data
fire_train3 <- fire_split3 %>%
                    training()


# Create testing data
fire_test3 <- fire_split3 %>%
                    testing()





fire_knn_fit <- fit(fire_knn_workflow, data = fire_train3)

test_features <- fire_test3 %>% select(temperature, isi) %>% data.frame()

nn1_pred3 <- predict(fire_knn_fit, test_features, type = "raw")

fire_results3 <- fire_test3 %>% 
  select(classes) %>% 
  bind_cols(predicted = nn1_pred3) %>% mutate(classes = as.factor(classes))


conf_mat(fire_results3, truth = classes, estimate = predicted)

```

```{r}
conf_mat(fire_results3, truth = classes, estimate = predicted)

```

```{r}
accuracy(fire_results3, truth = classes,
         estimate = predicted)

sens(fire_results3, truth = classes,
         estimate = predicted)

spec(fire_results3, truth = classes,
         estimate = predicted)

ppv(fire_results3, truth = classes,
         estimate = predicted)

```


*Answers:*

# Your Turn 3

a. Creating the recipe

```{r}
db_recipe <- recipe(
  diabetes ~  glucose + insulin, 
  data = db_raw
) %>%
  step_scale(all_predictors()) %>%
  step_center(all_predictors())
```


b. Create your model specification and use `tune()` as a placeholder for the number of neighbors

```{r}
knn_spec <- nearest_neighbor(
  weight_func = "rectangular", 
  neighbors = tune() #<<
) %>%
  set_engine("kknn") %>%
  set_mode("classification")
```


c. Split the `fire_train` data set into `v = 5` folds, stratified by `diabetes`

```{r}
db_vfold <- vfold_cv(db_raw, v = 5, strata = diabetes)
```

d. Create a grid of $K$ values, the number of neighbors and run 5-fold CV on the `k_vals` grid, storing four performance metrics

```{r}
k_vals <- tibble(neighbors = seq(from = 1, to = 15, by = 1))
```


```{r}
knn_fit <- workflow() %>%
  add_recipe(db_recipe) %>%
  add_model(knn_spec) %>%
  tune_grid(
    resamples = db_vfold, 
    grid = k_vals,
    metrics = metric_set(
      yardstick::recall, yardstick::ppv, 
      yardstick::accuracy, kap,
      roc_auc, sens, spec),
    control = control_resamples(
      save_pred = TRUE))
    
```


```{r}
cv_metrics <- collect_metrics(knn_fit)
cv_metrics
```

